---
title: "HWRIntermediate"
author: "Sarah Kell"
date: "January 17, 2018"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "C:/Users/Sarah/Desktop/CofC/Quant_Methods/RIntermediate")

setwd("C:/Users/Sarah/Desktop/CofC/Quant_Methods/RIntermediate")
```

*Can also type in 'i', 'j', 'x' etc. in code to see what is going on 

1.  Created a matrix/table that shows the averages for sepal length and width and petal length and width for each species in the dataframe 'iris'.

2. Describe using pseudo-code how output was calculated
```{r}
data("iris")
sp_ids = unique(iris$Species) #unique returns a vector, dataframe or array like x but with duplicate elements/rows removed. Pulled out the three species in iris and renamed them 'sp_ids' 

output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1) #creates matrix, 0 indicates matrix currently has no data, with rows of species names, with the number of columns within the dataframe 'iris' minus one column (to remove the species column)  
rownames(output) = sp_ids #assigns rownames in the matrix to the 3 species pulled out of the dataframe
colnames(output) = names(iris[ , -ncol(iris)]) #column names in matrix are column names in DF iris minus the last column specifically    

for(i in seq_along(sp_ids)) { 
  # Assigning i to a vector of the 3 species 
    iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    #Create a dataframe for each unique species with the column 'species'              removed 
    for(j in 1:(ncol(iris_sp))) {
      # Assign j to a vector that is 1 through the number of columns in                  the subset iris_sp (in this case 1:4) 
        x = 0 
        y = 0
        #create 2 objects, x and y, both defined as equal to 0
        if (nrow(iris_sp) > 0) {
          #if the number of rows in iris_sp is greater than zero do the                      following
            for(k in 1:nrow(iris_sp)) {
              #Assign k to a vector that is 1 through the number of rows                         in the subset iris_sp (1:50) 
                x = x + iris_sp[k, j] #define x as the sum of all numbers                          in iris_sp 
                y = y + 1 # y is equal to set to y +1 
                #print(paste('y=',y)) #add print (y) to see what is going                          on 
      
            }
            output[i, j] = x / y # The output will populate a matrix of x/y which              are the averages of the variable for each species  
        }
    }
}
output
```

3. Output could be renamed 'mean_values' 
x could be 'sum_parameters'
y could be 'row_length'     


4. Remove the for k loop and have x calculate the mean using the mean function. 
```{r}
sp_ids = unique(iris$Species) 

output = matrix(0, nrow=length(sp_ids), ncol=ncol(iris)-1)  
rownames(output) = sp_ids 
colnames(output) = names(iris[ , -ncol(iris)])    

for(i in seq_along(sp_ids)) { 
  iris_sp = subset(iris, subset=Species == sp_ids[i], select=-Species)
    for(j in 1:(ncol(iris_sp))) {
      if (nrow(iris_sp) > 0) {
        x = mean(iris_sp[, j]) 
        }
          output[i, j] = x 
    }
}
output
```

5. Creating loop 
```{r}
x<-c(1,2,3,4,5,6,7,8,9,10)

y<-vector("numeric", 10)
for(i in x) {
  y[i] <-sum(1:i)
}
y
```

6. Modify loop so values of y greater than 10 return NA
```{r}
x<-c(1,2,3,4,5,6,7,8,9,10)

y<-vector("numeric", 10)
for(i in x) {
  y[i] <-sum(1:i) 
    if (y[i] >10) {
      print("NA") }
    else {
      print(y[i]) 
    }
}
y
```

7. Turn loop into function
```{r}
x<-c(1,2,3,4,5,6,7,8,9,10)

sum_vector<-function (x) {
  y= NULL 
  for (i in x) {
    y[i] <-sum(1:i)
      if (y[i] >10) {
        print("NA") }
      else {
        print(y[i])
      }
  }
}
sum_vector(x)
```
